---
title: "covid 19"
author: "Lucille McEwen"
date: "October 20, 2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ There is no correlation between survival rate of covid-19 and sex of the patient. 
$H_A$ There is a correlation between survival rate of covid-19 and sex of the patient. 


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="Percentage Bar Chart")
```

By looking at the bar chart, it appears that the percentage of females with the virus is higher than the percentage of males with the virus. However, the males are dying at a higher percentage rate than females.  


  
## Numerical results
```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
``` 
  
The proportion of females who survived the virus is 56 percent whereas the proportion of males who survived the virus is 44 percent. Conversely, 36 percent of the females died as opposed to 64 percent of males. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds of dying if you're a male is 44/1411  
The odds of dying if you're a female is 25/1831  
The odds ratio is (44/1411)/(25/1831)   
The odds of dying of the virus are 2.3 to 1  



## Conclusions
The p-value is less than 0.05 so we reject the null hypothesis. There is a correlation between survival rates of covid-19 and sex of the patient. According to the data it appears that males are more likely to die from the virus than females. 
